Classical probability measures as pullback of quantum probability measures?

### 2022-10-29 pullback.tex
pullback.lyx Added some unreadable things... Comment out and commit whatever currently have so far...

%% LyX 2.3.6.1 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[11pt,english]{article}
\usepackage[T1,T2A]{fontenc}
\usepackage{verbatim}
\usepackage{booktabs}
\usepackage{textcomp}
\usepackage{mathrsfs}
\usepackage{enumitem}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{rotfloat}
\usepackage{setspace}
\usepackage[all]{xy}

@@

\usepackage{babel}
\usepackage[style=numeric-comp,backend=bibtex,maxbibnames=10,doi=false,isbn=false,url=false]{biblatex}
\addbibresource{prop.bib}
\begin{document}
\title{Classical probability measures as pullback of quantum probability
measures?}

@@

@@

A \emph{probability space} is a mathematical abstraction specifying
the necessary conditions for reasoning coherently about collections
of uncertain events \cite{Kolmogorov1950,544199,Griffiths2003,Grabisch2016}.
Although they can be used to describe an individual quantum experiment,
to describe a family of quantum experiments, we would like to glue
their probability spaces together to define a quantum probability
space. The glued quantum probability space is well-behaved since not
only we can define the expectation values, but the quantum probability
measure defined on the whole space can be simply induced by the Born
rule according to Gleason's theorem \cite{gleason1957,Redhead1987-REDINA,peres1995quantum,RichmanBridges1999,Hamhalter2013}.

@@

>>>>

%% LyX 2.3.6.1 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[11pt,ngerman,english]{article}
\usepackage[T1,T2A]{fontenc}
\usepackage{verbatim}
\usepackage{booktabs}
\usepackage{textcomp}
\usepackage{mathrsfs}
\usepackage{enumitem}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{rotfloat}
\usepackage{setspace}
\usepackage[all]{xy}

@@

\usepackage{babel}
\usepackage[style=numeric-comp,maxbibnames=10,doi=false,isbn=false,url=false]{biblatex}
\addbibresource{prop.bib}
\begin{document}
\title{Classical probability measures as pullback of quantum probability
measures?}

@@

\section{Introduction}

The means by which quantum computing\footnote{We use the phrase \textquotedblleft conventional quantum theory\textquotedblright{}
where necessary to distinguish the usual quantum theory and quantum
computing paradigm using (continuous) complex numbers from discrete
quantum theory. Alternative terminology in the literature includes
\textquotedblleft actual,\textquotedblright{} \textquotedblleft standard,\textquotedblright{}
and \textquotedblleft ordinary\textquotedblright{} quantum theory. } extends the capacity of classical computing frameworks deeply involves
both the laws of physics and the mathematical principles of computation.
Richard Feynman and \foreignlanguage{ngerman}{Rolf Landauer} \cite{Feynman1982Simulating,Landauer1996188},
among others, have strongly advocated the careful study of quantum
computation to understand its mechanisms and the source of its computational
features. To tease apart a number of subtle properties of quantum
computing, we previously replaced complex continuous numbers by finite
fields to study how uncomputable numbers affect quantum theory \cite{geometry2013,DQT2014},
and replaced real-valued probability measures by interval-valued probability
measures to study the effect of experimental imperfections and finite
precision measurements~\cite{THOS2018}. This time we try to decompose
a quantum probability space as a family of classical probability spaces
to see which part of quantum probability space has a classical counterpart,
to have a unified notation for classical and quantum probability for
this part, and to separate ``quantum-ness'' from the classical part
in a quantum probability description.\footnote{We might also put ``equivalence principle'' as one of the motivation.
Equivalence principle, gravitational and inertial forces are indistinguishable,
is the foundation to build relativity, and we are stating that in
some situation classical and quantum random process are indistinguishable...}

@@

A \emph{probability space} is a mathematical abstraction specifying
the necessary conditions for reasoning coherently about collections
of uncertain events \cite{Kolmogorov1950,544199,Griffiths2003,Grabisch2016}.
Since this abstraction doesn't specify the process to generate the
outcomes, we could use the same abstraction to describe both classical
and quantum processes, including tossing a coin and one Stern-Gerlach
type experiment \cite{:/content/aip/journal/jmp/21/1/10.1063/1.524312,peres1995quantum,schwinger2001quantum}.

Although a quantum process could be described by a probability space,
the probability spaces describing a family of quantum experiments
might contain some redundant information which could be used to glue
these probability spaces together to define a quantum probability
space. The glued quantum probability space is well-behaved since not
only we can define the expectation values, but the quantum probability
measure defined on the whole space can be simply induced by the Born
rule according to Gleason's theorem \cite{gleason1957,Redhead1987-REDINA,peres1995quantum,RichmanBridges1999,Hamhalter2013}.

@@

\footnote{Should find a ``simple'' example comparing the classical and quantum
process... Maybe Stern-Gerlach experiment with classical particle?}
